The Part-Time Parliament\hfill\insertframenumber/\inserttotalframenumber

### 2011-05-24 trunk/disciplinas/usi_algsem/paxosTalk/paxosTalk.tex
26 (woohoooooo) pages paxosTalk

@@

%==========================================================
\section{The Paxos Algorithm}
%==========================================================

@@

@@

%==========================================================
\section*{Summary}
%==========================================================

\begin{frame}
\frametitle<presentation>{Summary}

\begin{itemize}
  \item The \alert{first main message} of your talk in one or two lines.
\end{itemize}

% The following outlook is optional.
\vskip0pt plus.5fill
\begin{itemize}
  \item Outlook
  \begin{itemize}
    \item Something you haven't solved.
    \item Something else you haven't solved.
  \end{itemize}
\end{itemize}

@@

[1] Lamport, L. \textbf{The part-time parliament}, Technical Report 49, Systems Research Center, Digital Equipment Corp., 1989
\vspace{4 mm}

[2] Lamport, L. \textbf{Paxos Made Simple}, ACM SIGACT News, 2001
\vspace{4 mm}

[3] De Prisco, R., Lampson, B., Lynch, N. \textbf{Revisiting the Paxos Algorithm}, Distributed Algorithms, 1997
\vspace{4 mm}

[4] Lampson, B. \textbf{How to build a highly available system using consensus}, Distributed Algorithms, 1996
\vspace{4 mm}

[5] Castro, M., Liskov, B., \textbf{Practical Byzantine Fault Tolerance}, Operating Systems Design and Implementation, 1999

>>>>

  \vspace{3mm}

  Some people do use the described \textbf{Paxos} protocol: % 20 years after being first described?...

  \begin{itemize}
    \item Google uses Paxos in their Chubby distributed lock service
    \item IBM supposedly uses Paxos in their SAN Volume Controller product
    \item Microsoft uses Paxos in the Autopilot cluster management service
    \item WANDisco uses Paxos in their replication technology
    \item Keyspace uses Paxos as its basic replication primitive
  \end{itemize}

@@

%==========================================================
\section{About correctness}
%==========================================================

@@

%==========================================================
\section{The Synod Protocol}
%==========================================================

@@

  Phase 3)

  \vspace{6mm}
  Consistency is still maintained, and progress is still not guaranteed.

  \vspace{6mm}
  Anyway, it is impossible to guarantee termination of consensus in an asynchronous system [6]\ldots

%----------------------------------------------------------
\subsection{Complete protocol}
%----------------------------------------------------------

\begin{frame}
  \frametitle{Complete protocol}

  \begin {itemize}
    \item Same steps of the basic protocol

    \vspace{2mm}
    \item Each process performs phases 1b -- 3b as soon as they can

    \vspace{2mm}
    \item Some process must execute phase 1a
    \begin{itemize}
      \item Never initiating a ballot prevents termination
      \item However, too frequent initiation also does it %explicar cada ballot anulando o anterior, e sendo anulado antes de terminar
      \item A single process (leader) should be the only one initiating ballots %isso tb causaria a anulação de ballots anteriores de outros processos
    \end{itemize}

    \vspace{2mm}
    \item Knowledge of time, and of a maximum message delay $\Delta$, is needed
    \begin{itemize}
      \item Not exactly an asynchronous system anymore
    \end{itemize}
  \end{itemize}

\begin{frame}
  \frametitle{Complete protocol}

  \begin {itemize}
    \item A minimal execution of the basic protocol would take at most $5\Delta$
%     \begin {itemize}
%       \item the responses to message (1a) would arrive after $2\Delta$ (at most)
%       \item the responses to message (2a) would arrive after $2\Delta$ (at most)
%       \item the (3a) message would take at most $\Delta$
%     \end{itemize} 
    \vspace{2 mm}
    \item A leader would ensure progress, but that requires leader election     
    \begin {itemize}
      \item when a leader seems to have failed, a new one should be elected\\
      (the election takes time $T_{el}$)
      \item if not enough responses to (1a) or (2a) arrive after $2\Delta$, a ballot $B$ had started before $T_{el}$
      \begin{itemize}
        \item or too many failures or message losses occurred\ldots %which violates the assumptions for this protocol
      \end{itemize}
      %the other ballot may even have succeeded! 
      %maybe started by another process before T was fully elapsed
      \item in this case, a ballot with number $b' > B_{bal}$ may be started by the leader -- $NACK$s to (1a) or (2a) may have carried \bbal
      \item in the worst case, the leader receives a $NACK$ for (2a), $4\Delta$ after $T_{el}$
      \item as no other process initiated any ballot after $T_{el}$, the ballot numbered $b'$ would succeed $5\Delta$ after the $NACK$
    \end{itemize}
    \vspace{2 mm}
    \item The complete protocol, with leader election, terminates at most after $T_{el} + 9\Delta$
    \begin{itemize}
      \item if no message was lost, every correct process learns the decided value
    \end{itemize}

  \end{itemize}

%==========================================================
\section{The Paxos Protocol}
%==========================================================

%----------------------------------------------------------
\subsection{Running multiple consensus instances}
%----------------------------------------------------------

\begin{frame}
  \frametitle{The \textbf{Paxos} Protocol}
  \framesubtitle{Multiple consensus instances}

  \begin{itemize}
    \item Each instance has an id $I_{id}$ -- and there should be a leader %although, as seen, that wouldn't be really necessary for consistency
    \item Phases (1a) and (1b) can be executed for many of them --
    e.g. $p$ sends $NextBallot(b,n)$, for all instances $I$ with $I_{id} > n$
    \item The response from $q$ (1b) would contain a $LastVote(I_{id}, b, v)$ message for each instance $I : I_{id} > n$ that $q$ voted in
    \item Another process can send $NextBallot(b',n')$, with $b' > b$,\\
    and $p$'s (2a) messages would be ignored from then on
    \item This brings the protocol latency, on most cases, down to $3\Delta$
    \begin{itemize}
      \item no leader changes, enough correct processes and enough messages
    \end{itemize}
    \item If each process sent messages (2b) to every other one, we'd have $2\Delta$ %show how this is done on the board
    \item If instance $I^A$ was passed (phase (3b) concluded) before $I^B$ had a value proposed (phase (2a)), then $I^A_{id} < I^B_{id}$
    % ORDERING PROPERTY
    % the leader needed to know the result of previous instances before starting a new one (= atomic broadcast)
  \end{itemize}

%----------------------------------------------------------
\subsection{Relevance to computer science}
%----------------------------------------------------------

\begin{frame}
  \frametitle{The \textbf{Paxos} Protocol}
  \framesubtitle{Relevance to computer science}

  \begin{itemize}
    \item It may be used to implement state-machine replication
    \begin{itemize}
      \item A system (e.g. a database) may be implemented as a set of states and deterministic state changes
      \item The group of processes would be the set of replicas of the system
      \item Each state changed would be agreed upon by means of consensus (an ordered instance of Paxos)
      \item One of the replicas might be elected as leader, and client requests (state changes) would be sent to it
      \item As all replicas would have a consistent set of state changes, so would be their state
    \end{itemize} 
  \end{itemize}

  \vspace{2 mm}
  \begin{itemize}
    \item Many derivations from Paxos have been published since it was first described
    \begin{itemize}
      \item Cheap Paxos, Fast Paxos, Generalized Paxos, Multi-Paxos,\\
      Byzantine Paxos, Fast Byzantine Paxos, Fast Byzantine Multi-Paxos, Ring Paxos
    \end{itemize} 
  \end{itemize}

@@

[1] Lamport, L. \textbf{The part-time parliament}, Technical Report 49, Systems Research Center, Digital Equipment Corp., 1989

\vspace{4 mm}
[2] Lamport, L. \textbf{Paxos made simple}, ACM SIGACT News, 2001

\vspace{4 mm}
[3] De Prisco, R., Lampson, B., Lynch, N. \textbf{Revisiting the Paxos algorithm}, Distributed Algorithms, 1997

\vspace{4 mm}
[4] Lampson, B. \textbf{How to build a highly available system using consensus}, Distributed Algorithms, 1996

\vspace{4 mm}
[5] Castro, M., Liskov, B., \textbf{Practical Byzantine fault tolerance}, Operating Systems Design and Implementation, 1999

\vspace{4 mm}
[6] Fischer, M., Lynch, N., Paterson, M., \textbf{Impossibility of distributed consensus with one faulty process}, J. ACM 32 1 (Jan.), 374-382, 1985

\begin{center}
{\Huge \textbf{Thank you!}}

\vspace{5 mm}
{\footnotesize Questions?}
\end{center}
% 
%   \begin{itemize}
%     \item The paper presents the construction of a consensus protocol, based on three consistency conditions, from a preliminary one, to a complete one which ensures consistency \textbf{and} termination, as long as some message delay bound is known
%   \end{itemize}
% % 
% % The following outlook is optional.
% \vskip0pt plus.5fill
% \begin{itemize}
%   \item Outlook
%   \begin{itemize}
%     \item Something you haven't solved.
%     \item Something else you haven't solved.
%   \end{itemize}
% \end{itemize}
